Draft: Learning Laplacian from Binomial Signals

### 2022-05-26 Documents/paper/draft.Rmd
fix the figure citation

---
title: 'Draft: Learning Laplacian from Binomial Signals'
author: "Skip Moses"
date: "3/12/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
cls: IEEEtran.cls
bibliography: ref.bib
---

@@

## Introduction

@@

The methods for learning topological structure fall into the categories of regression like models @dong2016learning, @tugnait2021sparse, @pu2021learning, and @saboksayr2021accelerated, and machine learning techniques @pu2021kernel, @kalofolias2016learn, and  @venkitaraman2019predicting . In all of these models, signals are assumed to follow multivariate gaussian distributions, but there has been little exploration in learning a network from binomial signals.

@@

Because \ref{eq:dongmodel} is not jointly convex, Dong et al. employ an alternating minimization scheme to solve the problem. Initially, set $Y = X$ and we find a suitable $L$ by solving 
\begin{equation}
	\label{eq:dong:prob1}
	\begin{aligned}
		\min_L       & \phantom{..} \alpha \text{tr}(Y^TLY) + \beta ||L||_F^2, \\
		\textrm{s.t.}& \quad \text{tr}(L) = N,\\
					 & \quad L_{i,j} = L_{j,i} \leq 0, \phantom{..} i \neq j, \\
					 & \quad L\cdot \textbf{1} = \textbf{0}
	\end{aligned}
\end{equation}

Next, using $L$ from the first step we solve 
\begin{equation}
	\label{eq:dong:prob2}
	\begin{aligned}
		\min_Y \phantom{..} ||X-Y||_F^2 + \alpha \text{tr}(Y^TLY)
	\end{aligned}
\end{equation}

Both \ref{eq:dong:prob1} and \ref{eq:dong:prob2} can be cast as convex optimization problems. Specifically, \ref{eq:dong:prob1} can be solved with the method of alternating direction of multipliers (ADMM). The model is applied to both synthetic and real world data, and compared to a technique used in machine learning that is similar to sparse inverse covariance estimation of Gaussian Markov Random Field models. Four evaluation criteria were used  to evaluate the performance of the framework: $\textit{Precision, Recall, F-measure,}$ and $\textit{Normalized Mutual Information (NMI)}$. Kalofolias cites two weaknesses to this model: The use of the Frobenius norm is not amenable to interpretation, and it requires four constraints on the solution space @kalofolias2016learn.

@@

#### Signals on the graph and model specification
We consider a weighted undirected graph $G = (V, E)$, with the vertices set $V = {1, 2, \dots, N}$, and edge set $E$.  Let $\mathbf{A}$ denote the weighted adjacency matrix of $G$. In the case of weighted undirected graph, $\mathbf{A}$ is a square and symmetric matrix.

@@

##### Maximum likelihood for one stratum

@@

##### Optimization Program

@@

<hr />
\begin{align*}
&\textbf{Algorithm 1: Binary Signal Graph Learning} \\
&1) \textbf{ Input: } \text{ Input a signal } \mathbf{Y}.\\
&2) \textbf{ Output: } \text{ Output an estimated } \mathbf{A}.\\
&3) \textbf{ Initialization: } h_{i,0} \sim \mathcal{N}(0,1) \text{ for } i = 1,\ldots, N \\
&4) \textbf{ for } t = 1, \ldots, iter:\\
&5) \textbf{ Update } \mathbf{A} \textbf{ given } h:\\
&6) \,\,\,\, \textit{ Fix }  h \textit{ in Optimization Program and solve } \mathbf{A} \\
&7) \textbf{        Update } h \textbf{ given } \mathbf{A}:\\
&8) \,\,\,\, \textit{ Fix } \mathbf{A} \text{ in Optimization Program and solve h} \\
&9) \textbf{ end for }
\end{align*}
<hr /> 
\newpage
# Results
### Synthetic Data Generation

@@

### Graph Learning Results

@@

```{r EAG, echo=FALSE,out.width="80%", out.height="25%",fig.cap="Circular Embedding of Estimated graph. Note the overall weight of each edge is smaller than the edge weights of the ground truth.",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("pictures/est_graph.png"))
``` 
# Discussion
To our knowledge, this is the first work on learning graph topology from binary signals. We have developed a method to effectively learn the graph adjacency matrix from binary signals. This model can be easily extended for learning binomial signals in general. Incorporating parameter optimization techniques would be a natural improvement on this model. Moreover, it would be desirable to develop scalable algorithms that can be used to learn a topology for a large graph from signals. Recently, in @saboksayr2021accelerated the author shows fast proximal-gradient iterations can be applied to the framework given by @kalofolias2016learn improving the over-all runtime.

@@

## References
<div id="refs"></div>

>>>>

---
title: 'Draft: Learning Laplacian from Binomial Signals'
author: "Skip Moses"
date: "3/12/2022"
output:
  bookdown::pdf_document2: default
  html_document:
    df_print: paged
cls: IEEEtran.cls
bibliography: ref.bib
---

@@

# Introduction

@@

The methods for learning topological structure fall into the categories of regression like models @dong2016learning, @tugnait2021sparse, @pu2021learning, and @saboksayr2021accelerated, and machine learning techniques @pu2021kernel, @kalofolias2016learn, and  @venkitaraman2019predicting . In all of these models, signals are assumed to follow multivariate Gaussian distributions, but there has been little exploration in learning a network from binomial signals.

@@

Because (\ref{eq:dongmodel}) is not jointly convex, Dong et al. employ an alternating minimization scheme to solve the problem. Initially, set $Y = X$ and we find a suitable $L$ by solving 
\begin{equation}
	\label{eq:dongprob1}
	\begin{aligned}
		\min_L       & \phantom{..} \alpha \text{tr}(Y^TLY) + \beta ||L||_F^2, \\
		\textrm{s.t.}& \quad \text{tr}(L) = N,\\
					 & \quad L_{i,j} = L_{j,i} \leq 0, \phantom{..} i \neq j, \\
					 & \quad L\cdot \textbf{1} = \textbf{0}
	\end{aligned}
\end{equation}

Next, using $L$ from (\ref{eq:dongprob1}), we solve 
\begin{equation}
	\label{eq:dongprob2}
	\begin{aligned}
		\min_Y \phantom{..} ||X-Y||_F^2 + \alpha \text{tr}(Y^TLY)
	\end{aligned}
\end{equation}

Both (\ref{eq:dongprob1}) and (\ref{eq:dongprob2}) can be cast as convex optimization problems. Specifically, \ref{eq:dongprob1} can be solved with the method of alternating direction of multipliers (ADMM). The model is applied to both synthetic and real world data, and compared to a technique used in machine learning that is similar to sparse inverse covariance estimation of Gaussian Markov Random Field models. Four evaluation criteria were used  to evaluate the performance of the framework: $\textit{Precision, Recall, F-measure,}$ and $\textit{Normalized Mutual Information (NMI)}$. Kalofolias cites two weaknesses to this model: The use of the Frobenius norm is not amenable to interpretation, and it requires four constraints on the solution space @kalofolias2016learn.

@@

## Signals on the graph and model specification
We consider a weighted undirected graph $G = (V, E)$, with the vertices set $V = {1, 2, \dots, N}$, and edge set $E$.  Let $\mathbf{A}$ denote the weighted adjacency matrix of $G$. In the case of weighted undirected graph, $\mathbf{A}$ is a square and symmetric matrix.

@@

## Maximum likelihood for one stratum

@@

## Optimization Program

@@

<hr />
\begin{align*}
&\textbf{Algorithm 1: Binary Signal Graph Learning} \\
&1) \textbf{ Input: } \text{ Input a signal } \mathbf{Y}.\\
&2) \textbf{ Output: } \text{ Output an estimated } \mathbf{A}.\\
&3) \textbf{ Initialization: } h_{i,0} \sim \mathcal{N}(0,1) \text{ for } i = 1,\ldots, N \\
&4) \textbf{ for } t = 1, \ldots, iter:\\
&5) \textbf{ Update } \mathbf{A} \textbf{ given } h:\\
&6) \,\,\,\, \textit{ Fix }  h \textit{ in Optimization Program and solve } \mathbf{A} \\
&7) \textbf{        Update } h \textbf{ given } \mathbf{A}:\\
&8) \,\,\,\, \textit{ Fix } \mathbf{A} \text{ in Optimization Program and solve h} \\
&9) \textbf{ end for }
\end{align*}
<hr />

# Results
## Synthetic Data Generation

@@

## Graph Learning Results

@@

```{r EAG, echo=FALSE,out.width="80%", out.height="25%",fig.cap="Circular Embedding of Estimated graph. Note the overall weight of each edge is smaller than the edge weights of the ground truth.",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("pictures/est_graph.png"))
```

# Discussion
To our knowledge, this is the first work on learning graph topology from binary signals. We have developed a method to effectively learn the graph adjacency matrix from binary signals. This model can be easily extended for learning binomial signals in general. Incorporating parameter optimization techniques would be a natural improvement on this model. Moreover, it would be desirable to develop scalable algorithms that can be used to learn a topology for a large graph from signals. Recently, in @saboksayr2021accelerated the author shows fast proximal-gradient iterations can be applied to the framework given by @kalofolias2016learn improving the over-all runtime.

@@

<div id="refs"></div>
